Parameteric controls on responses to biogeochemical forcing in the CLM5

### 2018-11-19 main.tex

\authors{Rosie Fisher\affil{1},
William Wieder\affil{1},
Benjamin Sanderson\affil{2},
Charles Koven\affil{3},
Keith Oleson\affil{1},
Chonggang Xu\affil{4},
Ashehad Ali\affil{1},
Joshua Fisher\affil{5},
Mingjie Shi\affil{5},
Anthony Walker\affil{6},
Gordon Bonan\affil{1},
David Lawrence\affil{1}
}

@@

The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}). Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented by \cite{lawrence2018}, a description of the global Nitrogen cycling by Wieder et al. (in prep), the crop model by(Lombardozzi et al. in prep), the land use components by (Lawrence P et al. in prep), land surface hydrology by (Swenson et al. in prep) and plant hydraulics by (Kennedy et al. in prep).

CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4.0 and CLM4.5. The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing. Wieder et al. (in prep) report the global-scale implications of the full model update and note that the CLM5 responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.  In this paper, we describe the new components of the biogeochemical cycling, and investigate the features of the new model that control the model state and biogeochemical forcing responses in depth, using single-site experiments and parametric sensitivity analyses.

@@

One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}). The FUN model operates on the principle that for the alternative sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon. Further, the cost of each of each uptake pathway is variable in both time and space. FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to reduced growth and higher expenditure of C on N uptake (and reduces carbon use efficiency).  The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).

@@

Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant CN ratio:

@@

One issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}). The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity ((\cite{wieder2015} illustrated the impacts of uncertainty in this function).  FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compmared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}). Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation. Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground. In most cases, N fixation is preferred when soil N is low, and so and soil N builds up much quicker. In low productivity areas, however, carbon is required to fix N from a 'cold start' initial state, impacting the carbon use efficiency and survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.

\subsubsection{FLEXCN}
The results of the FACE model data comparison exercise illustrated that changes in the C:N ratio accounted for large changes in ecosystem carbon storage under elevated CO$_{2}$ (\cite{zaehle2014}, \cite{medlyn2015using}).  These changes were not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle. For this reason, we introduce a flexible tissue C:N ratio into the CLM5.

@@

The FUN model was originally conceived with a static tissue CN ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target CN ratio. In CLM5, one of the aims of the development process was to incorporate a varying CN ratio. To that end, it is necessary to modify the amount of Carbon expended on N uptake (and thus the amount of N required). The degree to which tissue CN content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments in light of limitations (including \cite{zaehle2010},\cite{ghimire2016}). Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from it's target N content. When N is expensive in the environment, less C is spent on uptake, increasing CN ratios where N is scarce. When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high. This new algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.

@@

Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible. We thus consider that an exploration of the CLM5 parameter space is also equivalent to an exploration of a large part of the relevant model structural space.

@@

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5. Here, we are focused on the recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization. Therefore, 
we focus on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle. The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.

Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model. However, we also report responses of the model state to perturbations. Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al. in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here. We expect that further focused analysed will be conducted by the CLM5 community, in tandem with analyses of coupled model experiments.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around a mean state, which allows more straightforward visualization and interpretation of the results. For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation). We recognize that a global parameter sensitivity analysis in this realm would also be of interest, and indeed have conducted many such analyses, but consider that assessment of the output of a global as well as a OAAT analysis of the entire domain of results presented here would reduce the opportunity for understanding the model parameter space in depth, as is the goal of our study.

@@

The parameters included in our sensitivity analysis are both detailed in (Table \ref{table_parameters}), and described in more depth below. \\

@@

2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, is the ratio of fine root biomass to leaf area. Unlike in previous versions of the CLM, the fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth. \\

3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody biomass, which in CLM5 is a constant fraction of leaf allocation. Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions. Woody biomass in CLM5 has no function and thus there is no cost to plants associated with low values of this parameter\\

4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5 we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The \emph(`n\_costs'}) parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake. \\

5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`fracfixers'}. This parameter is the fraction of the assimilation (gross primary productivity - autotropihic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM. \\

6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{`grperc'}, was revised substantially in CLM5 based on methodolicaly issues with the observation of growth respiration from Atkin O. (pers comm.) from 0.3 to 0.11. CLM5 also adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significnatly increases maintenance respiration costs. Here we test the impacts of this change on overall carbon budgets.\\

7) \textbf{Leaf carbon:nitrogen target}, \emph{leafcn}. CLM5 uses the concept of a 'target' leafcn ratio, around which flexibility is allowed, given the environmental costs of Nitrogen. LeafCN ratios are used also as input to the LUNA model to determine photosynthetic rate constants ($V_{c,max}$ and $J_{max}$) \\

8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}. This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, since it determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit (assuming no soil moisture lmitations).\\

@@

11,12,13) \textbf{Parameters a,b and c of the flexible CN model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in the CN technical note, determine the response of tissue CN ratios to depletion of N in both the environment and the plant tissues themselves.

@@

The Caxiuana site experienced substantial drought stress in these simulations (with $B_{tran,mn}$ of the default condition ~0.55), and little nutrient stress. Niwot Ridge indicated substantial amounts of drought ($B_{tran,mn}$ ~0.5), and N limitation, spending 27\% of carbon on nutrient acquisition. Harvard Forest site had a moderate drought stress ($B_{tran,mn}$~90\%), and moderate N limitation (~15\% of assimilation spent on uptake). The Metolius site had some drought stress ($B_{tran,mn}$ ~86\%) and very little nutrient stress.  At the sites with dought stress, it is evident that low values of the Medlyn parameter (which reduce stomatal conductance) relieved drought stress as expected from first principles, as do high C:N values, which reduce photosynthetic capacity, and thus also stomatal conductance. Lower root allocation increase drought stress, via reduction in root uptake capacity.

@@

Whether this impact on LAI is translated into an impact on GPP depends on the degree of saturation of productivity at the control LAI.

For GPP, the impact of SLA is always positive, given increasing LAI. There is a large impact at  but the degree of it's impact shifts between sites, as it integrates a trade-off between an increase in LAI and the deployment of thinner leaves (lower carbon content per unit area) and subsequent decreases the total N per unit area of leaf reducing photosynthetic capacity

The impact on NPP however, can be negative (e.g. Metolius), since the CLM5 can generate an under-storey of shaded leaves, and with it the possibility of those leaves being in negative carbon balance.  SLA driven increases in LAI can in principle also reduce NPP for respiratory reasons. At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the intrinsic trade-offs between costs and benefits of thinner leaves.

Impact of \emph{froot\_leaf}\\
Root:leaf ratio (\emph{froot\_leaf}) a direct effect on LAI, via changes in allocation fractions to different tissues, an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, and (newly introduced in CLM5) also affects the uptake of both nitrogen and water via it's impact on N uptake costs and hydraulic resistance terms respectively. The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.  Both types of effect are seen here, with the Niwot Ridge and Metolius site suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios, but the Harvard forest and Caxiuana sites showing increases.

@@

Decreases in stem:leaf ratio (stem\_leaf) below their default condition also caused a decline in leaf N per unit area (figure **e).  The causes of this behaviour are still not understood **

Impact of \emph{Ncosts}\\
The first-order impact of N$_{costs}$ is to increase the respiration losses associated with nitrogen uptake.  Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of N$_{costs}$ on the system state with the exception of the Niwot Ridge site, which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low. Other sites showed more muted responses in the vegetation state.

Impact of \emph{fracfixers}\\
Increasing \emph{Fracfixers} affects the maximum amount of carbon that plants can pay to symbiotically uptake Nitrogen. In all cases here, it has a relatively moderate and positive impact on the overall system state, in particular at the Niwot and Harvard forest sites. As is the case for N$_{costs}$, with high fixing capacity N\_uptake expenditure is lower, N uptake high, and NPP is moderately increased, resulting in increases in LAI and a slight response of GPP to that change.  These systems are only subject to transiently ramping CO$_{2}$, and are thus closer to equilibrium than the step-change fertilizations, suggesting that the ultimate state of the model N cycle might eventually be unaffected by N fixer fraction.

Impact of \emph{leafcn}\\
Changing the target leaf C:N ratio \emph{leafcn} over one standard deviation from the mean TRY values had substantial impacts on all the output variables.  As expected, higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.  In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge.

For NPP, the corresponding  decreases in respiration rates at high CN somewhat ameliorated this effect, particularly at Caxiuana, which has high ambient temperatures and low carbon use efficiency as a result.  At Niwot, an optimum leaf CN ratio was apparent in the range of the default values, indicating the possibility that other sites might be operating distant from their theoretical optima according to the model, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations. We do not find much evidence of saturation in the GPP/leafCN relationships, corresponding to a dominance of carboxylation-limited photosynthesis in these simulations.

In general, for a given change in NPP, the relative change in LAI was lower than for other parametric modification, reflecting the increased spending on nitrogen uptake inherent in the high $N_{area}$ simulations. Interestingly, despite a lower NPP for low CN values at Niwot Ridge, LAI values were higher, for reasons not yet understood ***

@@

Impact of \emph{medlyn\_slope}\\
The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties. At the Caxiuana site, the very lowest values caused a complete death of the ecosystem. It is unclear why this occured, since the impacts can be direct, on the carbon balance itself (not enough GPP to cover costs), but also on the leaf temperature (via surface energy balance issues) and hence the respiration rates. It is hard to discern the exact reasons, given the feedback inherent between all of these properties, but no impact on the carbon balance is found at intermediate levels.

At Niwot, a clear optimum was found in the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough. At Metolius and Harvard forest, only lower value affected the photosynthetic rates, indicating co-limitation by another process, and the absence of impacts of excessive plant transpiration on the system (note that \cite{kennedy} and \cite{dagon} discuss surface energy and plant hydrodynamic functioning of the model in much greater depth).

Impact of \emph{lmr\_intercept}\\
The variance in the reported intercepts in \cite{atkin2016} also have important consequences for ecosystem productivity, in line with the uncertainties in the growth respiration parameterization. For the high leaf area simulations (now shown **), the impacts of maintenance respiration parameterization were much higher, reflecting the higher amount of respiring tissue in those simulations. At Caxiuana, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.

@@

Impact of \emph{cn\_flex\_a}\\
\emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake.  (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details. Thus, low values of \emph{cn\_flex\_a}  results in reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.   For the range investigated here, this parameter had limited impact on the system state.

Impact of \emph{cn\_flex\_b}\\
\emph{cn\_flex\_b} also increases the amount of carbon spent on N uptake when costs are high, but has limited impacts on the system state here. At low values, there is some indication at Niwot Ridge that it might increase LAI and vegetation carbon, indicating a decrease in the N uptake expenditure and in leaf N concentrations.  The limited impact at other sites suggests that Ncosts are rarely high enough to trigger reductions in N expenditure (notwithstanding the point the this algorithm is a placeholder for a more evidence-based depiction of stoichiometric flexibility).

>>>>

\authors{Rosie Fisher\affil{1},
William Wieder\affil{1},
Benjamin Sanderson\affil{2},
Charles Koven\affil{3},
Keith Oleson\affil{1},
Chonggang Xu\affil{4},
Ashehad Ali\affil{1},
Joshua Fisher\affil{5},
Mingjie Shi\affil{5},
Anthony Walker\affil{6},
Gordon Bonan\affil{1},
David M. Lawrence\affil{1}
}

@@

The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}). Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented by \cite{lawrence2018}, a description of the global Nitrogen cycling by Wieder et al. (in prep), the crop model by(Lombardozzi et al. in prep), the land use components by (Lawrence P et al. in prep), and plant hydraulics by (Kennedy et al., 2018).

CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4.0 and CLM4.5. The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing. Wieder et al. (in prep) report the global-scale carbon cycle implications of the full model update and note that the CLM5 carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors.  In this paper, we describe the new components of the biogeochemical cycling, and investigate the features of the new model that control the model state and biogeochemical forcing responses in depth, using single-site experiments and parametric sensitivity analyses.

@@

One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}). The FUN model operates on the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon. Further, the cost of each uptake pathway is variable in both time and space. FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to reduced growth and higher expenditure of C on N uptake (and reduces carbon use efficiency).  The central element of FUN is a set of simultaneous equations which assume, firstly, that the total carbon available (after maintenance respiration and replenishment of stores are accounted for), `$C_{avail}$', is either used for plant growth ($C_{growth}$) or for nitrogen uptake, ($C_{nuptake}$).

@@

Further, the nitrogen acquired from the environment ($N_{uptake}$) must equal the amount spent on N divided by it's average cost ($N_{cost}$), and also the amount of carbon deployed for growth divided by the plant CN ratio:

@@

One issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}). The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity ((\cite{wieder2015} illustrated the impacts of uncertainty in this function).  FUN considers fixation rates to be an emergent property of plant assimilation rates, the relative costs of environmental N acquisition compared to fixation, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}). Fixation is preferred when the costs of `active' N acquisition from soil are higher than that of fixation. Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground. In most cases, N fixation is preferred when soil N is low, and so soil N builds up quickly. In low productivity areas, however, carbon is required to fix N from a 'cold start' initial state, impacting the carbon use efficiency and survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.

\subsubsection{FLEXCN}
The results of FACE model-data comparison exercises illustrated that changes in the C:N ratio accounted for large changes in ecosystem carbon storage under elevated CO$_{2}$ (\cite{zaehle2014}, \cite{medlyn2015using}).  These changes were not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle. For this reason, we introduce a flexible tissue C:N ratio into the CLM5.

@@

The FUN model was originally conceived with a static tissue C:N ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target C:N ratio. In CLM5, one of the aims of the development process was to incorporate a varying C:N ratio. To that end, it is necessary to modify the amount of Carbon expended on N uptake (and thus the amount of N required). The degree to which tissue C:N content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments under N limited conditions (including \cite{zaehle2010},\cite{ghimire2016}). Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from it's target N content. When N is expensive in the environment, less C is spent on uptake, increasing C:N ratios where N is scarce. When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high. This new algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.

@@

Because of the increased number of prognostic processes with the CLM5, many alternative representations of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible. We thus consider that an exploration of the CLM5 parameter space is also equivalent to an exploration of a large part of the relevant model structural space.

@@

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the CLM5. For this manuscript, we focus on parameters included in recent developments in the vegetation model, and how they might impact upon the responses to CO$_{2}$ and N fertilization. Therefore, 
we focus on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle. The choice of parameters used here is informed primarily by the model structure, as well as many iterative and informal investigations of the CLM5 model during the development process.

Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model. However, we also report responses of the model state to perturbations. Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al. in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here. We expect that further focused analyses will be conducted by the CLM5 community, in tandem with analyses of coupled model experiments.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around a mean state, which allows more straightforward visualization and interpretation of the results. For each parameter, for each site, we conduct four simulations, with two simulations each using parameter values above and  below the default value (as well as the default simulation). We recognize that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses, but consider that assessment of the output of a global as well as a OAAT analysis of the entire domain of results presented here would reduce the opportunity for understanding the model parameter space in depth, as is the goal of our study.

@@

The parameters included in our sensitivity analysis are both detailed in Table \ref{table_parameters}, and described in more depth below. \\

@@

2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, is the ratio of fine root biomass to leaf area. Unlike in previous versions of the CLM, the fine root biomass in CLM5 affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth. \\

3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody biomass, which in CLM5 is a constant fraction of leaf allocation. Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions. Woody biomass in CLM5 has no function and thus there is no cost to plants associated with low values of this parameter.\\

4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The \emph(`n\_costs'}) parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake. \\

5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`fracfixers'}. This parameter is the fraction of the assimilation (gross primary productivity - autotrophic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM. \\

6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{`grperc'}, was revised substantially in CLM5 to reflect new estimates that account for methodological issues with the observation of growth respiration (Atkin O., pers comm.) from 0.3 to 0.11. CLM5 also adopts a new leaf maintenance respiration scheme, based on \cite{atkin2015}, which significantly increases maintenance respiration costs. Here we test the impacts of this change on overall carbon budgets.\\

7) \textbf{Leaf carbon:nitrogen target}, \emph{leafcn}. CLM5 uses the concept of a 'target' leaf C:N ratio, around which flexibility is allowed, given the environmental costs of Nitrogen. Leaf C:N ratios are used also as input to the LUNA model to determine photosynthetic rate constants ($V_{c,max}$ and $J_{max}$). \\

8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}. This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependent), is critically linked to the response of ecosystems to CO$_{2}$, since it determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit (assuming no soil moisture limitations).\\

@@

11,12,13) \textbf{Parameters a,b and c of the flexible C:N model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves.

@@

The Caxiuana site experienced substantial drought stress in these simulations (with $B_{tran,mn}$ of the default condition ~0.55), and little nutrient stress. Niwot Ridge indicated substantial amounts of drought ($B_{tran,mn}$ ~0.5), and N limitation, spending 27\% of carbon on nutrient acquisition. Harvard Forest site had a moderate drought stress ($B_{tran,mn}$~90\%), and moderate N limitation (~15\% of assimilation spent on uptake). The Metolius site had some drought stress ($B_{tran,mn}$ ~86\%) and very little nutrient stress.

These differences in the initial conditions at the sites have implications for the parameteric responses, as documented for each parameter in turn in the section below.

@@

For GPP, the impact of SLA is always positive, given increasing LAI, and despite higher LAI causing lower N per unit area and leaves, reducing photosynthetic capacity. The impact of SLA on NPP can however be negative (e.g. Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance. At Niwot Ridge, an optimal SLA is observed with respect to NPP, again reflecting the assimilation/respiration cost trade-off.

Impact of \emph{froot\_leaf}\\
Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in allocation fractions to different tissues, 2) an indirect effect on NPP, given the modified maintenance costs of the resultant fine root biomass pools, 3) affects the uptake of water via root biomass and 4) of nutrients via root-mediated n uptake costs.  The first two impacts increase NPP with \emph{froot\_leaf} and the latter two decrease it.  Both types of effect are seen here, with the Niwot Ridge and Metolius sites (which have the lowest productivity and LAI, and thus presumably very low root biomass in the first place) suffering very large reductions in GPP, NPP, and LAI at low \emph{froot\_leaf} ratios.  Harvard forest and Caxiuana both show increases in all variables.  In Figure \ref{btran state}, we can see that low \emph{froot\_leaf} is a primary control over drought stress at the low productivity sites.

@@

Decreases in stem:leaf ratio (stem\_leaf) below their default condition also caused a decline in leaf N per unit area, potentially on account of the higher demand for N uptake (illustrated clearly for Niwot Ridge in Figure \ref{btran state}) when the LAI is higher.

Impact of \emph{Ncosts}\\
The first-order impact of $N_{costs}$ is to increase the respiration losses associated with nitrogen uptake.  Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of $N_{costs}$ on the system state with the exception of the Niwot Ridge site, where N limitation was highest and which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low. Other sites showed more muted responses in the vegetation state.

Impact of \emph{fracfixers}\\
Increasing \emph{Fracfixers} affects the maximum amount of carbon that plants can pay to symbiotically uptake Nitrogen. In all cases here, it has a relatively moderate and positive impact on the overall system state, in particular at the Niwot and Harvard forest sites. As is the case for $N_{costs}$, with high fixing capacity N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that change.

Impact of \emph{leafcn}\\
Changing the target leaf C:N ratio \emph{leafcn} over one standard deviation from the mean TRY values had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.  In particular, the higher C:N ratios had leafCN and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).    We do not find much evidence of saturation in the GPP/leafCN relationships, corresponding to a dominance of carboxylation-limited photosynthesis in these simulations, in line with findings in \cite{lawrence} that V${_c,max}$ numbers might be low compared to observations.

For NPP, the corresponding decreases in respiration rates caused by high \emph{leafcn}, and reduced expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuana, which has high ambient temperatures and low carbon use efficiency as a result. At Niwot, an optimum leaf CN ratio was apparent. For NPP this was around the default value, but for LAI it was 0.5s.d. lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.

@@

Impact of \emph{medlyn\_slope}\\
The change in the stomatal efficiency parameter of the Medlyn 2011 stomatal conductance model most directly impacts GPP, with cascading impacts on all downstream properties (including on leaf temperature and water stress). At the Caxiuana site (Figure \ref{CAX state}), the very lowest values caused a complete death of the ecosystem. In figure 
\ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses). It is worth noting that these same simulations typically had extremely low photosynthesis, particularly at Caxiuana and Metolius.

At Niwot, a clear optimum was found in the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough. At Metolius and Harvard forest, only the lower values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance. (note that \cite{kennedy} and \cite{dagon} discuss surface energy and plant hydrodynamic functioning of the model in greater depth).

Impact of \emph{lmr\_intercept}\\
The variance in the reported intercepts in \cite{atkin2016} also have important consequences for ecosystem productivity, in line with the uncertainties in the growth respiration parameterization. At Caxiuana, the warmest site here, the \emph{lmr\_intercept} was one of the most important contributing factors to the NPP, LAI and vegetation carbon states.

@@

Impact of \emph{cn\_flex\_a}\\
\emph{cn\_flex\_a}  is the intercept above which an increase in the cost of N will reduce expenditure on N uptake (see \href{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html#modifications-to-allow-variation-in-c-n-ratios} for details). Thus, low values of \emph{cn\_flex\_a}  results in reductions in expenditure on N uptake when N is too expensive, and therefore on lower tissue N content, and vice-versa.   For the range investigated here, this parameter had limited impact on the system state.

Impact of \emph{cn\_flex\_b}\\
Increasing \emph{cn\_flex\_b} also increases the amount of carbon spent on N uptake when costs are high, and also has limited impacts on the system state here.

Impact of \emph{cn\_flex\_c}\\
 \emph{cn\_flex\_c} affects the degree to which plants attempt to maintain tissue CN values near to their target when environmental costs are varying. For the state impacts here, the role of this parameter was limited, but for the step-change fertilizations it was much more important, so we will discuss it's role further in the section below.
